---
title: "DATA607Project1"
author: "Erick Hadi"
date: "2024-10-06"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DATA 607 Project 1 

In this project, you’re given a text file with chess tournament results where the information has some structure. Your
job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database)
with the following information for all of the players:
Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents

## Load the data
```{r}
library(tidyverse)
library(dplyr)
library(tidyr)
```

```{r}
gitlink <- "https://raw.githubusercontent.com/ErickH1/DATA607Project1/refs/heads/main/tournamentinfo.txt"
chess_raw_data <- readLines(gitlink)
head(chess_raw_data, 7)
```
## Data Processing 
Inserting text data into a matrix to make data capturing and processing easier.
```{r}
chess_data_matrix <- matrix(unlist(chess_raw_data), byrow=TRUE)
matrix_1 <- chess_data_matrix[seq(5,length(chess_data_matrix),3)]
matrix_2 <- chess_data_matrix[seq(6,length(chess_data_matrix),3)]

head(chess_data_matrix,10)
head(matrix_1)
head(matrix_2)
```
## Extracting Chess Data
Using Regex and string manipulation to extract relevant information into vectors.
```{r}
ID <- as.numeric(str_extract(matrix_1, '\\d+'))

Name <- str_trim(str_extract(str_extract(matrix_1, '[A-z].{1,32}'), '.+\\s{2,}'))

State <- str_extract(matrix_2, '[A-Z]{2}') 

Total_Points <- as.numeric(str_extract(matrix_1, '\\d+\\.\\d'))

Pre_Rating <- as.numeric(str_extract(str_extract(matrix_2, 'R:.{8,}-'), '\\d{1,4}'))

Rounds <- str_extract_all(matrix_1, '[A-Z]\\s{2,}\\d+')
Rounds <- str_extract_all(Rounds, '\\d+')
```
## Calculate Avg Opponent Rating
Calculating avg opponent rating using pre rating and rounds vectors. Instead of for loops utilized sapply.
```{r}
Avg_Opp_Pre_Rating <- sapply(Rounds, function(x) round(mean(Pre_Rating[as.numeric(x)]), 0))
Avg_Opp_Pre_Rating
```
## Inserting Extracted Data Into Chess Data Frame
```{r}
chess_data <- data.frame(ID,Name,State,Total_Points,Pre_Rating,Avg_Opp_Pre_Rating)
head(chess_data)
```

## Export Data to CSV 
```{r}
write.csv(chess_data, "chesstournamentinfo.csv")
```